---
title: "Activity 02- Bivariate Viz"
---

```{r}
# Load data
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")

# Check it out
head(elections)
```
Exercise 0: Review

```{r}
ggplot(elections, aes(x=winner_20)) +
  geom_bar()

ggplot(elections, aes(repub_pct_20)) +
  geom_histogram(fill="blue")
```

Exercise 1 Quantitative vs Quantitative Intuition Check

```{r}
ggplot(elections, aes(x=repub_pct_16, y=repub_pct_20)) +
  geom_point()
```

Exercise 2: 2 Quantitiative Variables
```{r}
# Set up the plotting frame
# How does this differ than the frame for our histogram of repub_pct_20 alone?
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16))

# Add a layer of points for each county
# Take note of the geom!
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point()

# Change the shape of the points
# What happens if you change the shape to another number?
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(shape = 4)

# YOU TRY: Modify the code to make the points "orange"
# NOTE: Try to anticipate if "color" or "fill" will be useful here. Then try both.
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(color = "orange")
```

Exercise 3: Reflect
Summarize the relationship between the Republican support in 2020 and 2016. Be sure to comment on:

the strength of the relationship (weak/moderate/strong)
the direction of the relationship (positive/negative)
outliers (in what state do counties deviate from the national trend? Any ideas why this might be the case?)

Republican support in 2020 and 2016 have a strong positive correlation. There are 
