---
title: "Activity 01- Data Viz"
---

```{r, message=FALSE}
# Load the package
library(tidyverse)

# Import data
hikes <- read.csv("https://mac-stat.github.io/data/high_peaks.csv")
```

```{r}
head(hikes)
```

What features would we like a visualization of the categorical difficulty rating variable to capture?
- A bar chart would be good for a categorical variable such as difficulty.

What about a visualization of the quantitative elevation variable?
- A histogram would be good for elevation variable.


```{r}
# Use the ggplot function
ggplot(hikes, aes(x = rating))

# Adding geom_bar() indicates what kind of graph to create
ggplot(hikes, aes(x = rating)) +
  geom_bar()

# Changes the axis labels
ggplot(hikes, aes(x = rating)) +
  geom_bar() +
  labs(x = "Rating", y = "Number of hikes")

# Changes the color of the bar chart
ggplot(hikes, aes(x = rating)) +
  geom_bar(fill = "blue") +
  labs(x = "Rating", y = "Number of hikes")

# Adds an orange outline
ggplot(hikes, aes(x = rating)) +
  geom_bar(color = "orange", fill = "blue") +
  labs(x = "Rating", y = "Number of hikes")

# Changes the theme of the graph, making the background white
ggplot(hikes, aes(x = rating)) +
  geom_bar(color = "orange", fill = "blue")  +
  labs(x = "Rating", y = "Number of hikes") +
  theme_minimal()
```

Part a
Reflect on the ggplot() code.

What’s the purpose of the +? When do we use it?
- the + indicates that another function is being added to the ggplot function.

We added the bars using geom_bar()? Why “geom”?
- It indicates that the graph is geometric

What does labs() stand for?
- labs() stands for labels

What’s the difference between color and fill?
- color changes the color of the border, fill changes the color of the entire bar.

Part b
In general, bar charts allow us to examine the following properties of a categorical variable:
observed categories: What categories did we observe?
- difficulty

variability between categories: Are observations evenly spread out among the categories, or are some categories more common than others?
- moderate is by far the most common category

We must then translate this information into the context of our analysis, here hikes in the Adirondacks. Summarize below what you learned from the bar chart, in context.

Part c
Is there anything you don’t like about this barplot? For example: check out the x-axis again.
- the x-axis is in alphabetical order not order of difficulty.


Exercise 6
```{r}
ggplot(hikes, aes(x = elevation)) +
  geom_bar()
```
This is NOT an effective visualization because it creates a new bar for every unique elevation, making the bars small and difficult to read as very few peaks have the exact same elevation.

Exercise 8
```{r}
ggplot(hikes, aes(x = elevation)) +
  geom_histogram()
```
Exercise 9
```{r}
# Adds a histogram plot
ggplot(hikes, aes(x = elevation)) +
  geom_histogram()

# Adds white border to bars in plot
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white") 

# Makes the bars blue
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", fill = "blue") 

# Adds axis labels
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white") +
  labs(x = "Elevation (feet)", y = "Number of hikes")

# Changes the range of each bar to 1000
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", binwidth = 1000) +
  labs(x = "Elevation (feet)", y = "Number of hikes")

# Changes the range of each bar to 2
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", binwidth = 5) +
  labs(x = "Elevation (feet)", y = "Number of hikes")

# Changes the range of each bar to 200
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", binwidth = 200) +
  labs(x = "Elevation (feet)", y = "Number of hikes")
```

Exercise 11
```{r}
ggplot(hikes, aes(x = elevation)) +
  geom_density(color = "blue", fill = "orange")
```

ggplot template:
```{r}
#ggplot(___, aes(x = ___)) + 
  #geom___(color = "___", fill = "___") + 
  #labs(x = "___", y = "___")
```


PRACTICE
```{r}
# Data on students in this class
survey <- read.csv("https://ajohns24.github.io/data/112/about_us_2024.csv")

# World Cup data
world_cup <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-11-29/worldcups.csv")
```

```{r, survey plots}
ggplot(data=survey, aes(x = minutes_to_campus)) + 
  geom_histogram(color = "purple", fill = "navy")

ggplot(data=world_cup, aes(x=host)) +
  geom_bar(fill="green")
```

