---
title: "Journal"
---

# 3. Data Viz

Data Viz: displaying data in such a way that we can glean more information about the data, such as patterns, typical outcomes, and distribution.

### Components of graphs:
- frame (sets up the grid and axes)
- layer (adds the geometric elements that make up the graph, called geom_ in ggplot)
- scales (changes color/size/shape/etc)
- faceting (splits up the data )
- theme (finer scale adjustments like font size)

### For one variable viz:
- categorical: bar plot
- numerical: histogram/density plot

ggplot syntax:
```{r, eval=FALSE}
ggplot(data = ___, aes(x = ___)) + 
  geom___(color = "___", fill = "___") + 
  labs(x = "___", y = "___")
```

### When discussing data visualizations, you typically look at:
- typical outcome
- variability + range
- shape
- outliers


# 4. Bivariate Viz

Another way to look at the data is to look at relationships within the data. There are
- Repsonse variables (dependent variables)
- Predictors (independent variables)

### What plots to use:
- 2 quantitative: scatterplot
- 2 categorical: stacked bar/facet wrapped bar plot
- 1 quantitative, 1 categorical: violin plot, boxplot, stacked density plots.

# 5. Multivariate Viz

Which chart to use:

- 3 cat: bar (x, color, facet)
- 2 cat, 1 num: histogram/density (color, facet, x), box, violin (x, facet, y)
- 1 cat, 2 num: scatter (facet/color/shape, x, y), heatmap (y, x, shade), star (y, size, )
- 3 num: scatter (x, y, size/shade)


# 6. Spatial Viz

- Point map: a map that specific locations are plotted on (ex: favorite places in the twin cities).
- Contour map: plotting density/distribution on a map rather than individual points
- Choropleth Maps: plotting outcomes in different regions (ex: percentage of democratic votes in each state)

Leaflet: a package that can be used to make interactive maps, commonly point maps
common functions for leaflet:

- leaflet() (makes the leaflet)
- addTiles() (makes the base map)
- addMarkers()
- addPolygons()
- print() (need to print to see the leaflet)

ggplot is typically better for making chloropleth maps than leaflet.


# 7. Effective Viz

Effective visualizations are important in order to accurately portray your data in a way that people can understand. 

Visualizations can be good, ugly, bad, or wrong. 

- Good: viz is clear, easy to understand, accessible, and helpful.
- Ugly: nothing is incorrect about the viz, but the overall aesthetic is ugly
- Bad: the viz is “unclear, confusing, overly complicated, or deceiving”
- Wrong: viz is incorrectly displaying the data and/or the trends

In order to have effective viz you need to:

- have clear, concise titles/axes
- include a figure caption
- include an alt caption, for screen readers
- avoid using overly similar colors whenever possible
- add context
- make data anonymous when possible


# 8. Wrangling Verbs

tidyverse has many helpful functions that allow us to wrangle the data into a way that we can meaningfully create data viz.

99.99% of data will not be entirely clean and ready to use. You will have to clean and reformat the data in order to use it.

Some key functions to use are: 



| verb        | action                                            |
|:------------|:--------------------------------------------------|
| `arrange`   | **arrange** the *rows* according to some *column* |
| `filter`    | **filter** out or obtain a subset of the *rows*   |
| `select`    | **select** a subset of *columns*                  |
| `mutate`    | **mutate** or create a *column*                   |
| `summarize` | calculate a numerical **summary** of a *column*   |
| `group_by`  | **group** the *rows* by a specified *column*      |



# 9. Wrangling + Dates

Order of the functions matter. If you want to arrange by one column but then only include 2 different columns in your final data frame, you need to do arrange first before selecting the columns or else you will get an error.

Key syntax for filtering:

| symbol           | meaning                   |
|:-----------------|:--------------------------|
| `==`               | equal to                  |
| `!=`               | not equal to              |
| `>`               | greater than              |
| `>=`              | greater than or equal to  |
| `<`               | less than                 |
| `<=`              | less than or equal to     |
| `%in% c(***, ***)` | a list of multiple values |

summarize ex:

``` {r, eval = FALSE}
penguins |> filter(species == “Chinstrap”) |> group_by(sex) |> summarize(min = min(body_mass_g), max = max(body_mass_g)) |> mutate(range = max - min)
```

# 10. Reshaping Data

Two kinds of reshaped data:

- Aggregate: more general info about observations gained using functions like group_by() with summarize(), no info about individual observations
- Raw data: original data that is just reshaped to be able to analyze.


Pivoting:

pivot wider: "widens" data by expanding two columns into multiple. One column provides the new column names and the other provides the values.

```{r, eval = FALSE}
data |> 
  pivot_wider(names_from = col1, values_from = col2)
```


pivot longer: "lengthens" the data by collapsing several columns into two. Column names move to a new names_to column and values to a new values_to column.

```{r, eval = FALSE}
data |> 
  pivot_longer(cols = c("columns we want"), names_to = "name of new names column", values_to = "name of new values column")
```


# 11. Joining Data

Joining functions allow us to either filter one dataset based on another or to join two datasets togeher. There are two different types of joining functions:

- mutating joins: left_join(), inner_join() and full_join()
- filtering joins: semi_join(), anti_join()


Mutating Joins:

> left_join: join two datasets together, keep all rows in the left dataset and add in matching columns from right dataset. Automatically uses a common column to join datasets, but you can specify which one/ones you want to use as the join key.

```{r, eval = FALSE}
left_data %>% 
  left_join(right_data, c("L_col" = "R_col")) # indicating that those two are the join key
```

> inner_join: keeps only the columns that are in both the left and right datasets.

```{r, eval = FALSE}
left_data %>% 
  inner_join(right_data)
```


> full_join: keeps the columns in the right AND/OR the left datasets

```{r, eval = FALSE}
left_data %>% 
  full_join(right_data)
```


Filtering Joins:

> semi_join: keep all rows in left that match right

```{r, eval = FALSE}
left_data %>% 
  semi_join(right_data)
```

> anti_join: keep all rows in left that do not match in right

```{r, eval = FALSE}
left_data %>% 
  anti_join(right_data)
```


# 12. Factors

When working with character strings, you can either use regex to wrangle them or convert them into factors.

Some of the factor wrangling functions:

-   functions for changing the **order** of factor levels
    -   `fct_relevel()` = *manually* reorder levels
    -   `fct_reorder()` = reorder levels according to values of another *variable*
    -   `fct_infreq()` = order levels from highest to lowest frequency
    -   `fct_rev()` = reverse the current order
-   functions for changing the **labels** or values of factor levels
    -   `fct_recode()` = *manually* change levels
    -   `fct_lump()` = *group together* least common levels


# 13. Strings

Common string modifying/locating functions:

| function            | arguments                 | returns           |
|:--------------------|:--------------------------|:------------------|
| `str_replace()`     | `x, pattern, replacement` | a modified string |
| `str_replace_all()` | `x, pattern, replacement` | a modified string |
| `str_to_lower()`    | `x`                       | a modified string |
| `str_sub()`         | `x, start, end`           | a modified string |
| `str_length()`      | `x`                       | a number          |
| `str_detect()`      | `x, pattern`              | TRUE/FALSE        |

Summary of what each function does:

- str_replace(x, pattern, replacement) finds the first part of x that matches the pattern and replaces it with replacement

- str_replace_all(x, pattern, replacement) finds all instances in x that matches the pattern and replaces it with replacement

- str_to_lower(x) converts all upper case letters in x to lower case

- str_sub(x, start, end) only keeps a subset of characters in x, from start (a number indexing the first letter to keep) to end (a number indexing the last letter to keep)

- str_length(x) records the number of characters in x

- str_detect(x, pattern) is TRUE if x contains the given pattern and FALSE otherwise


# 15. Data Import

Rather than using absolute filepaths that will not work on other people's computers, use relative filepaths

```{r, eval = FALSE}
imdb_temp <- read_csv("../data/imdb_5000_messy.csv")

#the .. means exit current folder
```


